Are Delayed Issues Harder to Resolve? Revisiting Cost-to-Fix of Defects throughout the Lifecycle

### 2016-07-06 phaseDelay4Emse.tex
typo

The DIE theory is worth examination since it has been used as the basis for decision-making in software engineering. For example,  Basili and Boehm comment that, since the 1980s,  this effect
 \begin{quote}
``...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes''~\cite{boehm01}.
\end{quote}
 \begin{figure}[!b]  
\begin{center}
\includegraphics[,width=4in]{img/b81.png}
 \end{center}
 \caption{A widely-recreated chart of the DIE effect. Adapted from Boehm'81~\cite{Boehm81}. }\label{fig:b81}
 \end{figure}

@@

\begin{figure}[!t]
% \begin{center}\includegraphics[width=4in]{img/boehm_cost-to-fix.png}\end{center}
\begin{center}\includegraphics[width=4.5in]{img/boehm81.png}\end{center}
 \caption{Historical cost-to-fix curve. Adapted from~\cite{Boehm81}, p. 40.}\label{fig:cost-to-fix}
 \end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=2.5in]{img/defect_type_frequency.png}
\end{center}
\caption{Relative frequencies of these defect types seen in our TSP data.}\label{fig:dtypes}.
\end{figure}
In our TSP data,  the relative frequencies of these defect types are shown in \fig{dtypes}. Around a quarter of the fixes were simple documentation changes.
That said, 75\% of the changes are quote  elaborate;   e.g. fixes to function necessitates a careful reflection of the purpose of the code.

@@

\subsubsection{Definition: development phase} \label{development_phase}
\begin{figure}[!b]  
\begin{center}
\includegraphics[width=1.6in]{img/wfall.png}  
\end{center}
\caption{Phases of our data.
Abbreviations: 
{\em Before}= before development; 
{\em Reqts}	  = requirements; 
{\em HLD}	  = high-level design; 
{\em IntTest} = Integration testing (with code from others); 
{\em SysTest} = system test (e.g. load stress tests); 
{\em AcceptTest}  = acceptance testing (with users); 
{\em review}        = private activity; 
{\em inspect}        = group activity.}
\label{fig:waterfall}
\end{figure}

@@

\begin{figure}[!t] 
\begin{center}
\includegraphics[height=1.5in]{img/number_of_projects_per_org.png}
\end{center} 
\caption{Number of projects per development organization.}
\label{fig:number of projects per organization}
\end{figure}

@@

 \begin{figure}[ht]
\scriptsize
\centering
\begin{tabular}{lrrrrrrrl}
                                                                   & N   & Min & Q1   & Median & Q3    & Max   & Mean   &  Distribution \\\hline
Team size                                                          & 171 & 1   & 4    & 6      & 10    & 36    &  7.8    &  \includegraphics[width=15mm]{img/team_size.png} \\ 
Duration {[}days{]}                                                & 171 & 7   & 33   & 61     & 118   & 1918  & 107    & \includegraphics[width=15mm]{img/duration.png} \\
Added \& Modified LOC                                                 & 117 & 2   & 1125 & 4201   & 13092 & 88394 & 10259  & \includegraphics[width=15mm]{img/LOC.png} \\
Components                                                         & 171 & 8   & 26   & 49     & 107   & 4170  & 116.2  & \includegraphics[width=15mm]{img/components.png} \\
Year                                                               & 171 &     & 2011 & 2012   & 2013  & 2014  & 2011.9 & \includegraphics[width=15mm]{img/year.png}
\end{tabular}

@@

\begin{figure}[!t] 
\vspace{0.5cm}
\begin{center}
\includegraphics[height=2.5in]{img/fix-phase-dist.pdf}
\end{center} 
\caption{Distribution of defects found and fixed by phase.}
\label{fig:fix-phase-dist}
\end{figure}

@@

\begin{figure} 
% \renewcommand{\baselinestretch}{0.8}
% \small
\begin{center}
\input{tex/fig13}
%
% \begin{tabular}{l@{~~~}r|r@{~}r|r@{~}lr@{~}l} 
% \renewcommand{\baselinestretch}{0.5}
% \multicolumn{2}{l|}{Phase}&\multicolumn{2}{c|}{Percentile}&\multicolumn{4}{c}{Ratio w.r.t. to phase where issue was injected}\\\hline
%  injected&removed&50th&95th&\multicolumn{2}{c|}{50th}&\multicolumn{2}{c}{95th}\\\hline
% \\
% Before&DesignInspect&30&97&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rulree{10mm}{2mm}} \\
% &CodeReview&28&94&0.93&\rule{9mm}{2mm}&0.96&\textcolor{red}{\rule{10mm}{2mm}} \\
% &CodeInspect&45&154&1.5&\rule{15mm}{2mm}&1.58&\textcolor{red}{\rule{16mm}{2mm}} \\
% &UnitTest&77&187&2.56&\rule{26mm}{2mm}&1.92&\textcolor{red}{\rule{19mm}{2mm}} \\
% &IntTest&65&424&2.16&\rule{22mm}{2mm}&4.37&\textcolor{red}{\rule{35mm}{2mm}} \\
% &SysTest&56&164&1.86&\rule{19mm}{2mm}&1.69&\textcolor{red}{\rule{17mm}{2mm}} \\\hline\\
% Planning&ReqtsReview&35&99&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &DesignInspect&24&81&0.68&\rule{7mm}{2mm}&0.81&\textcolor{red}{\rule{8mm}{2mm}} \\
% &UnitTest&33&238&0.94&\rule{9mm}{2mm}&2.4&\textcolor{red}{\rule{24mm}{2mm}} \\\hline\\
% Reqts&ReqtsReview&92&240&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &ReqtsInspect&93&243&1.01&\rule{10mm}{2mm}&1.01&\textcolor{red}{\rule{10mm}{2mm}} \\
% &DesignReview&33&119&0.35&\rule{4mm}{2mm}&0.49&\textcolor{red}{\rule{5mm}{2mm}} \\
% &DesignInspect&49&137&0.53&\rule{5mm}{2mm}&0.57&\textcolor{red}{\rule{6mm}{2mm}} \\
% &CodeInspect&38&156&0.41&\rule{4mm}{2mm}&0.65&\textcolor{red}{\rule{7mm}{2mm}} \\
% &UnitTest&49&180&0.53&\rule{5mm}{2mm}&0.75&\textcolor{red}{\rule{8mm}{2mm}} \\
% &IntTest&67&300&0.72&\rule{7mm}{2mm}&1.25&\textcolor{red}{\rule{13mm}{2mm}} \\
% &SysTest&122&459&1.32&\rule{13mm}{2mm}&1.91&\textcolor{red}{\rule{19mm}{2mm}} \\\hline\\
% Design&DesignReview&103&331&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &DesignInspect&101&272&0.98&\rule{10mm}{2mm}&0.82&\textcolor{red}{\rule{8mm}{2mm}} \\
% &CodeReview&66&215&0.64&\rule{6mm}{2mm}&0.64&\textcolor{red}{\rule{6mm}{2mm}} \\
% &CodeInspect&79&259&0.76&\rule{8mm}{2mm}&0.78&\textcolor{red}{\rule{8mm}{2mm}} \\
% &UnitTest&119&383&1.15&\rule{12mm}{2mm}&1.15&\textcolor{red}{\rule{12mm}{2mm}} \\
% &IntTest&88&337&0.85&\rule{9mm}{2mm}&1.01&\textcolor{red}{\rule{10mm}{2mm}} \\
% &SysTest&80&336&0.77&\rule{8mm}{2mm}&1.01&\textcolor{red}{\rule{10mm}{2mm}} \\
% &AcceptTest&45&120&0.43&\rule{4mm}{2mm}&0.36&\textcolor{red}{\rule{4mm}{2mm}} \\\hline\\
% Code&CodeInspect&113&337&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &CodeReview&115&300&1.01&\rule{10mm}{2mm}&0.89&\textcolor{red}{\rule{9mm}{2mm}} \\
% &UnitTest&133&434&1.17&\rule{12mm}{2mm}&1.28&\textcolor{red}{\rule{13mm}{2mm}} \\
% &IntTest&125&473&1.1&\rule{11mm}{2mm}&1.4&\textcolor{red}{\rule{14mm}{2mm}} \\
% &SysTest&95&383&0.84&\rule{8mm}{2mm}&1.13&\textcolor{red}{\rule{11mm}{2mm}} \\
% &AcceptTest&65&360&0.57&\rule{6mm}{2mm}&1.06&\textcolor{red}{\rule{11mm}{2mm}} \\

@@

\begin{figure}[!t]
% \renewcommand{\baselinestretch}{0.8}
% \small
\begin{center}
\input{tex/figlast}
\end{center}
\caption{90th percentile tiles to resolve issues in the SEI TSP data. Same format as \fig{raw} (but here we look at 90th percentile
outliers while   \fig{raw} explored the central tendencies of the data). The conclusion from this figure is the same
as \fig{raw}: even when considering the 90th percentile outliers, there is no consistent patterns of 
exponential growthe in the effort to fix issues delayed later in the lifecycle.}\label{fig:last}
\end{figure}

@@

\input{tex/responses}

\end{document}

>>>>

The DIE theory is worth examination since it has been used as the basis for decision-making in software engineering. For example,  Basili and Boehm comment that, since the 1980s,  this effect
 \begin{quote}
``...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes''~\cite{boehm01}.
\end{quote}
 \begin{figure}[!b]  
\begin{center}
\includegraphics[,width=4in]{b81.png}
 \end{center}
 \caption{A widely-recreated chart of the DIE effect. Adapted from Boehm'81~\cite{Boehm81}. }\label{fig:b81}
 \end{figure}

@@

\begin{figure}[!t]
% \begin{center}\includegraphics[width=4in]{boehm_cost-to-fix.png}\end{center}
\begin{center}\includegraphics[width=4.5in]{boehm81.png}\end{center}
 \caption{Historical cost-to-fix curve. Adapted from~\cite{Boehm81}, p. 40.}\label{fig:cost-to-fix}
 \end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=2.5in]{defect_type_frequency.png}
\end{center}
\caption{Relative frequencies of these defect types seen in our TSP data.}\label{fig:dtypes}.
\end{figure}
In our TSP data,  the relative frequencies of these defect types are shown in \fig{dtypes}. Around a quarter of the fixes were simple documentation changes.
That said, 75\% of the changes are quote  elaborate;   e.g. fixes to function necessitates a careful reflection of the purpose of the code.

@@

\subsubsection{Definition: development phase} \label{development_phase}
\begin{figure}[!b]  
\begin{center}
\includegraphics[width=1.6in]{wfall.png}  
\end{center}
\caption{Phases of our data.
Abbreviations: 
{\em Before}= before development; 
{\em Reqts}	  = requirements; 
{\em HLD}	  = high-level design; 
{\em IntTest} = Integration testing (with code from others); 
{\em SysTest} = system test (e.g. load stress tests); 
{\em AcceptTest}  = acceptance testing (with users); 
{\em review}        = private activity; 
{\em inspect}        = group activity.}
\label{fig:waterfall}
\end{figure}

@@

\begin{figure}[!t] 
\begin{center}
\includegraphics[height=1.5in]{number_of_projects_per_org.png}
\end{center} 
\caption{Number of projects per development organization.}
\label{fig:number of projects per organization}
\end{figure}

@@

 \begin{figure}[ht]
\scriptsize
\centering
\begin{tabular}{lrrrrrrrl}
                                                                   & N   & Min & Q1   & Median & Q3    & Max   & Mean   &  Distribution \\\hline
Team size                                                          & 171 & 1   & 4    & 6      & 10    & 36    &  7.8    &  \includegraphics[width=15mm]{team_size.png} \\ 
Duration {[}days{]}                                                & 171 & 7   & 33   & 61     & 118   & 1918  & 107    & \includegraphics[width=15mm]{duration.png} \\
Added \& Modified LOC                                                 & 117 & 2   & 1125 & 4201   & 13092 & 88394 & 10259  & \includegraphics[width=15mm]{LOC.png} \\
Components                                                         & 171 & 8   & 26   & 49     & 107   & 4170  & 116.2  & \includegraphics[width=15mm]{components.png} \\
Year                                                               & 171 &     & 2011 & 2012   & 2013  & 2014  & 2011.9 & \includegraphics[width=15mm]{year.png}
\end{tabular}

@@

\begin{figure}[!t] 
\vspace{0.5cm}
\begin{center}
\includegraphics[height=2.5in]{fix-phase-dist.pdf}
\end{center} 
\caption{Distribution of defects found and fixed by phase.}
\label{fig:fix-phase-dist}
\end{figure}

@@

\begin{figure} 
% \renewcommand{\baselinestretch}{0.8}
% \small
\begin{center}

%\resizebox{\columnwidth}{!}{
\scriptsize
\begin{tabular}{c|lr|rr|rl}
& \multicolumn{2}{c|}{ } & \multicolumn{2}{c|}{Percentiles} & \multicolumn{2}{c}{~ } \\ 
  & \multicolumn{2}{c|}{ } & \multicolumn{2}{c|}{(units = } & \multicolumn{2}{c}{Growth with respect to earliest phase  } \\ 
   & \multicolumn{2}{c|}{Phase} & \multicolumn{2}{c|}{  minutes)} & \multicolumn{2}{c}{(unitless  ratios of two time values) } \\\hline

  rank & injected & removed & 50th &   IQR & \multicolumn{2}{c}{50th percentile growth}  \\ 
\hline

\multicolumn{7}{c}{~}  \\
1 &  Before   
      &   DesignInspect   & 10 &  14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   CodeReview      & 8 &  14 & 0.80 & \textcolor{black}{\rule{8mm}{2mm}}  \\
1 &   &   CodeInspect     & 10 &  16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   UnitTest        & 12 &   21 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &   &   IntTest         & 15 &   31 & 1.50 & \textcolor{black}{\rule{15mm}{2mm}}  \\
1 &   &   SysTest         & 11 &   22 & 1.10 & \textcolor{black}{\rule{12mm}{2mm}}  \\  
\hline\multicolumn{7}{c}{~}  \\
1 &  Planning     &   ReqtsReview     & 8 &  14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &               &   DesignInspect   & 11 &  13 & 1.38 & \textcolor{black}{\rule{13mm}{2mm}} \\
2 &               &   UnitTest        & 24 &   25 & 3.00 & \textcolor{black}{\rule{30mm}{2mm}} \\
\hline\multicolumn{7}{c}{~}  \\
1 &  Reqts   &   ReqtsReview     & 13 &  20 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} \\
1 &          &   ReqtsInspect    & 12 &   18 & 0.92 & \textcolor{black}{\rule{9mm}{2mm}}  \\
1 &          &   DesignReview    & 10 &   14 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}} \\
1 &          &   DesignInspect   & 9 &   15 & 0.69 & \textcolor{black}{\rule{6mm}{2mm}}   \\
1 &          &   CodeInspect     & 13 &  24 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} \\
1 &          &   UnitTest        & 10 &  17 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}}  \\
1 &          &   IntTest         & 33 &   42 & 2.54 & \textcolor{black}{\rule{25mm}{2mm}}  \\
1 &          &   SysTest         & 24 &  108 & 1.85 & \textcolor{black}{\rule{18mm}{2mm}}  \\
\hline\multicolumn{7}{c}{~}  \\
1 & Design   &   DesignReview    & 11 &   16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   DesignInspect   & 8 &   12 & 0.73 & \textcolor{black}{\rule{7mm}{2mm}}  \\
1 &   &   CodeReview      & 10 &   18 & 0.91 & \textcolor{black}{\rule{9mm}{2mm}}  \\
1 &   &   CodeInspect     & 9 &  14 & 0.82 & \textcolor{black}{\rule{8mm}{2mm}}  \\
1 &   &   UnitTest        & 11 &   18 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   IntTest         & 17 &   31 & 1.55 & \textcolor{black}{\rule{15mm}{2mm}}  \\
1 &   &   AcceptTest      & 14 &   19 & 1.27 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &   &   SysTest         & 13 &   18 & 1.18 & \textcolor{black}{\rule{12mm}{2mm}}  \\
\hline\multicolumn{7}{c}{~}  \\
1 &  Code   &   CodeReview    & 10 &  16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &    &   CodeInspect   & 10 &   15 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   UnitTest      & 12 &   20 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &    &   IntTest       & 14 &   25 & 1.40 & \textcolor{black}{\rule{14mm}{2mm}}  \\
1 &    &   AcceptTest    & 16 &   25 & 1.60 & \textcolor{black}{\rule{16mm}{2mm}}  \\
1 &   &   SysTest       & 13 &   20 & 1.30 & \textcolor{black}{\rule{13mm}{2mm}}  \\

\end{tabular}
%}
%
% \begin{tabular}{l@{~~~}r|r@{~}r|r@{~}lr@{~}l} 
% \renewcommand{\baselinestretch}{0.5}
% \multicolumn{2}{l|}{Phase}&\multicolumn{2}{c|}{Percentile}&\multicolumn{4}{c}{Ratio w.r.t. to phase where issue was injected}\\\hline
%  injected&removed&50th&95th&\multicolumn{2}{c|}{50th}&\multicolumn{2}{c}{95th}\\\hline
% \\
% Before&DesignInspect&30&97&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rulree{10mm}{2mm}} \\
% &CodeReview&28&94&0.93&\rule{9mm}{2mm}&0.96&\textcolor{red}{\rule{10mm}{2mm}} \\
% &CodeInspect&45&154&1.5&\rule{15mm}{2mm}&1.58&\textcolor{red}{\rule{16mm}{2mm}} \\
% &UnitTest&77&187&2.56&\rule{26mm}{2mm}&1.92&\textcolor{red}{\rule{19mm}{2mm}} \\
% &IntTest&65&424&2.16&\rule{22mm}{2mm}&4.37&\textcolor{red}{\rule{35mm}{2mm}} \\
% &SysTest&56&164&1.86&\rule{19mm}{2mm}&1.69&\textcolor{red}{\rule{17mm}{2mm}} \\\hline\\
% Planning&ReqtsReview&35&99&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &DesignInspect&24&81&0.68&\rule{7mm}{2mm}&0.81&\textcolor{red}{\rule{8mm}{2mm}} \\
% &UnitTest&33&238&0.94&\rule{9mm}{2mm}&2.4&\textcolor{red}{\rule{24mm}{2mm}} \\\hline\\
% Reqts&ReqtsReview&92&240&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &ReqtsInspect&93&243&1.01&\rule{10mm}{2mm}&1.01&\textcolor{red}{\rule{10mm}{2mm}} \\
% &DesignReview&33&119&0.35&\rule{4mm}{2mm}&0.49&\textcolor{red}{\rule{5mm}{2mm}} \\
% &DesignInspect&49&137&0.53&\rule{5mm}{2mm}&0.57&\textcolor{red}{\rule{6mm}{2mm}} \\
% &CodeInspect&38&156&0.41&\rule{4mm}{2mm}&0.65&\textcolor{red}{\rule{7mm}{2mm}} \\
% &UnitTest&49&180&0.53&\rule{5mm}{2mm}&0.75&\textcolor{red}{\rule{8mm}{2mm}} \\
% &IntTest&67&300&0.72&\rule{7mm}{2mm}&1.25&\textcolor{red}{\rule{13mm}{2mm}} \\
% &SysTest&122&459&1.32&\rule{13mm}{2mm}&1.91&\textcolor{red}{\rule{19mm}{2mm}} \\\hline\\
% Design&DesignReview&103&331&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &DesignInspect&101&272&0.98&\rule{10mm}{2mm}&0.82&\textcolor{red}{\rule{8mm}{2mm}} \\
% &CodeReview&66&215&0.64&\rule{6mm}{2mm}&0.64&\textcolor{red}{\rule{6mm}{2mm}} \\
% &CodeInspect&79&259&0.76&\rule{8mm}{2mm}&0.78&\textcolor{red}{\rule{8mm}{2mm}} \\
% &UnitTest&119&383&1.15&\rule{12mm}{2mm}&1.15&\textcolor{red}{\rule{12mm}{2mm}} \\
% &IntTest&88&337&0.85&\rule{9mm}{2mm}&1.01&\textcolor{red}{\rule{10mm}{2mm}} \\
% &SysTest&80&336&0.77&\rule{8mm}{2mm}&1.01&\textcolor{red}{\rule{10mm}{2mm}} \\
% &AcceptTest&45&120&0.43&\rule{4mm}{2mm}&0.36&\textcolor{red}{\rule{4mm}{2mm}} \\\hline\\
% Code&CodeInspect&113&337&1&\rule{10mm}{2mm}&1&\textcolor{red}{\rule{10mm}{2mm}} \\
% &CodeReview&115&300&1.01&\rule{10mm}{2mm}&0.89&\textcolor{red}{\rule{9mm}{2mm}} \\
% &UnitTest&133&434&1.17&\rule{12mm}{2mm}&1.28&\textcolor{red}{\rule{13mm}{2mm}} \\
% &IntTest&125&473&1.1&\rule{11mm}{2mm}&1.4&\textcolor{red}{\rule{14mm}{2mm}} \\
% &SysTest&95&383&0.84&\rule{8mm}{2mm}&1.13&\textcolor{red}{\rule{11mm}{2mm}} \\
% &AcceptTest&65&360&0.57&\rule{6mm}{2mm}&1.06&\textcolor{red}{\rule{11mm}{2mm}} \\

@@

\begin{figure}[!t]
% \renewcommand{\baselinestretch}{0.8}
% \small
\begin{center}
%\resizebox{\columnwidth}{!}{
\scriptsize
\begin{tabular}{c|lr|r|rl}
  & \multicolumn{2}{c|}{ } &  Percentiles & \multicolumn{2}{c}{Growth with respect to earliest phase  } \\ 
   & \multicolumn{2}{c|}{Phase} &  (units= miniutes) & \multicolumn{2}{c}{(unitless ratios of two time values) } \\\hline

  rank & injected & removed &   90th   &  \multicolumn{2}{c}{90th percentile growth} \\ 
\hline\multicolumn{6}{c}{~}  \\
1 &  Before   &   DesignInspect     & 32  &  1.00 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &   &   CodeReview      &   31   &    0.97 & \textcolor{black}{\rule{9mm}{2mm}}\\
1 &   &   CodeInspect     &   32   &    1.00 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &   &   UnitTest        &   45  &    1.41 & \textcolor{black}{\rule{14mm}{2mm}}\\
1 &   &   IntTest         &   63  &   1.97 & \textcolor{black}{\rule{19mm}{2mm}}\\
1 &   &   SysTest         &   46   &   1.44 & \textcolor{black}{\rule{14mm}{2mm}}\\  
\hline\multicolumn{6}{c}{~}  \\
1 &  Planning     &   ReqtsReview       & 35   &   1.00 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &               &   DesignInspect     & 31  &   0.89 & \textcolor{black}{\rule{8mm}{2mm}}\\
2 &               &   UnitTest          & 53 &   1.51 & \textcolor{black}{\rule{15mm}{2mm}}\\
\hline\multicolumn{6}{c}{~}  \\
1 &  Reqts   &   ReqtsReview      & 42 &    1.00 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &          &   ReqtsInspect      & 40 &     0.95 & \textcolor{black}{\rule{9mm}{2mm}}\\
1 &          &   DesignReview      & 34 &    0.81 & \textcolor{black}{\rule{8mm}{2mm}}\\
1 &          &   DesignInspect     & 38 &     0.90 & \textcolor{black}{\rule{9mm}{2mm}}\\
1 &          &   CodeInspect      & 45 &   1.07 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &          &   UnitTest          & 40 &   0.95 & \textcolor{black}{\rule{9mm}{2mm}}\\
1 &          &   IntTest          & 95 &      2.26 & \textcolor{black}{\rule{22mm}{2mm}}\\
1 &          &   SysTest          & 126 &    3.00 & \textcolor{black}{\rule{30mm}{2mm}}\\
\hline\multicolumn{6}{c}{~}  \\
1 & Design   &   DesignReview     & 37 &   1.00 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &          &   DesignInspect    & 28 &    0.76 & \textcolor{black}{\rule{7mm}{2mm}}\\
1 &   &   CodeReview        & 40 &     1.08 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &   &   CodeInspect      & 33 &     0.89 & \textcolor{black}{\rule{8mm}{2mm}}\\
1 &   &   UnitTest          & 41 &     1.11 & \textcolor{black}{\rule{12mm}{2mm}}\\
1 &   &   IntTest          & 75 &    2.03 & \textcolor{black}{\rule{20mm}{2mm}}\\
1 &   &   AcceptTest       & 44 &     1.19 & \textcolor{black}{\rule{12mm}{2mm}}\\
1 &   &   SysTest           & 40 &     1.08 & \textcolor{black}{\rule{10mm}{2mm}}\\
\hline\multicolumn{6}{c}{~}  \\
1 &  Code   &   CodeReview      & 35    & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}\\
1 &    &   CodeInspect    & 32   & 0.91 & \textcolor{black}{\rule{9mm}{2mm}}\\
1 &   &   UnitTest      & 45  & 1.29 & \textcolor{black}{\rule{12mm}{2mm}}\\
1 &    &   IntTest         & 58   & 1.66 & \textcolor{black}{\rule{16mm}{2mm}}\\
1 &    &   AcceptTest      & 60 &   1.71 & \textcolor{black}{\rule{17mm}{2mm}}\\
1 &   &   SysTest        & 47 &    1.34 & \textcolor{black}{\rule{13mm}{2mm}}\\

\end{tabular}
%}
\end{center}
\caption{90th percentile tiles to resolve issues in the SEI TSP data. Same format as \fig{raw} (but here we look at 90th percentile
outliers while   \fig{raw} explored the central tendencies of the data). The conclusion from this figure is the same
as \fig{raw}: even when considering the 90th percentile outliers, there is no consistent patterns of 
exponential growthe in the effort to fix issues delayed later in the lifecycle.}\label{fig:last}
\end{figure}

@@

 \section*{Response to Reviewers}

{\bf Editor's comments}

Dear authors,

Three knowledgeable reviewers have scrutinized the manuscript
and found it interesting and relevant,  although in need for
certain improvements. Therefore, the authors are
 invited to submit a revised version for a new review, where
 the reviewers' comments
are addressed.

\begin{quote} {\em Thank you for the careful reviews. Using
those comments we have extensively revised and extended this draft.
Specifically, the old version was 20 pages and this version is 29 pages (and the new
material all comes from the clarification and correction points offered by the reviewers).
We hope that  you will find this draft greatly improved.}
\end{quote}

Particularly, this holds for:

1) the connection to activities of theory building in ESE
(reviewers 1).

\begin{quote}{\em This new draft now maps our work into the SE literature on theory building. This new material appears in the introduction as well as throughout the document.}\end{quote}

2) clarification about the data origin and
characteristics (reviewer 1).

\begin{quote}{\em
This new draft contains much more detail on the origin and characteristics of our data (see \tion{data-collection} and \tion{data_character}).}\end{quote}

3)  use of statistical analysis (reviewers 2 and 3)

\begin{quote}{\em All our results are now augmented with statistical significance tests as well as effect size tests (see \tion{stats} and \tion{171projects}). Those tests strengthen the overall message of this paper-- that there is no significant
phase delay effect.}\end{quote}

- writing style, that would benefit form being more
precise (reviewer 3)

\begin{quote}{\em We have carefully read those comments and have made numerous changes
as directed by reviewer3.}\end{quote}

{\bf Reviewer \#1's Comments}

The paper investigates whether a widely
accepted "rule" in software engineering holds by looking at
a large set of empirical data. The research topic is highly
relevant and the results are interesting.   The discussion
on reassessing old truism fits the scope of the journal very
well.  The language in the paper is excellent, and the
structure is (although non-standard) easy to follow.

\begin{quote}{\em Thank you for those kind comments.}\end{quote}

It is not easy for the reader to understand what parts
of the "life of an issue" the authors claim are affected by
DIE, and it is also not too easy to understand what data the
authors have available.

\begin{quote}{\em  The introduction now includes an example of DIE  (see Fig 1).
Also there are now more precise  definitions on how we measure this effect in \tion{data-collection}}\end{quote}

 The authors do not relate their reassessment of an
``old truism'' to the growing set of papers on theory building
in software engineering. With access to such large amounts
of empirical evidence, it should be possible to take some
steps toward improved DIE theories.

\begin{quote}{\em We now map our work to prior publications on theory-building in software engineering, as well as theory falsification in general. That mapping appears in numerous comments through out the text (using the references offered by this reviewer) with the bulk of the comments being top of p2 and the conclusion.

Notably, we now introduce the DIE as an example of such a theory (in the introduction), and explain that this is done on the basis of it providing a description of a Software development phenomenon accompanied by a rationale that explains why such behavior is expected. While we introduce in the conclusions several rationales that can explain why the effect called for by the theory was not observed in our dataset, we feel that future work is needed to collect data in support or contradiction of these explanations before a new theory can be formulated with confidence. }\end{quote}

The different projects studied are all grouped together. I
believe studying clusters of development contexts would be
highly interesting, but unfortunately the current
characterization of projects is inadequate.

\begin{quote}{\em  We agree with this comment and so took two actions for this draft-- one that was effective and one that was ineffective.

First, this draft now contains extensive notes on the different contexts of the
projects studied here. Those notes are offered in \tion{data_character} and \tion{171projects}.

Second (and this was the ineffective step), we  searched the categories   listed in \tion{data_character} and \tion{171projects}
in order to check if any stratification contained the delayed issue effect. 
That study   failed to find a stratification
containing an exponential expansion in the time to fix issues. The reason for this was  the small size
of those stratifications  exacerbated the large IQR's seen in this data\footnote{Recall that in a sorted list of numbers,
the inter-quartile range, or IQR, is difference between the 
  75th to 25th percentile value.}:
\bi
\item
Our 171 projects   stratifies into subsets of varying sizes. Sorting those sizes, we get a list
comprising of 17,12 then numerous much smaller stratifications.  Reasoning over such small samples
is problematic in the general case and, in the case of our data, it is ever more problematic due to
the large IQRS of the data:
\item
To see these large IQRs,  please consider the 50th and IQR columns of Figure~\ref{fig:raw} Note that most of our IQRs were larger than tan the 50th percentile; i.e. software
data exhibits large variances-- where are exacerbated by the smaller samples seen in the stratifications.
\ei
Our conclusion from exploring the stratification's is, given the currently available data, we cannot check
for a DIE effect in subsets of this data.

Note that the above notes are now added to the paper at the end of Section 6.6.
 }\end{quote}

DETAILED REVIEW: The authors should expand on what they
include in the ``delayed issue effect'' (DIE), in particular
in relation to standard milestones in the life of an issue
and the corresponding timespans.  I'm particularly
interested in whether all "indirect" issue resolution steps
are covered, e.g., change impact analysis, updating
documentation, and regression testing. The effort involved
in these steps clearly depends on the development context of
the specific project. The short a-c) list on page 14:50
suggest that the issues studied are restricted to minor
fixes, i.e., no updates to architecture, changed hardware,
recertification, updated user manuals etc.

\begin{quote} {\em
We now provide a more explicit definition of what constitutes a defect (our study of issues is limited to defects) and the ``defect lifecycle'' phases we measure in \tion{data-collection}. These definitions greatly expand on the now-deleted list on 14:50 in teh previous version. In TSP, ``a defect is any change to a product, after its construction, that is necessary to make the product correct''. This includes both minor issues (typography) and severe issues (architecture changes, requirements errors). Time on a defect includes a) time to investigate/analyze a defect once discovered, b) time to craft and implement a fix, and c) time to verify and close the defect.

Regarding the concern that we may only be reporting trivial defects, we have added \fig{dtypes}, which shows the distribution of defect types in our data. It is true that around a quarter of the fixes were documentation changes, but the other 75\% are function-altering behavior in the system. 
}\end{quote}

Section 6.3 describes the data, but some aspects
should be further explained.

\begin{quote} {\em
At your suggestion, we have added extensive notes on the nature of our data--
see \tion{data_character} and \tion{171projects}.
}\end{quote}

Logging interruption time
sounds very useful, but I wonder how carefully the
developers actually did this 
 it must be really hard to keep up the discipline required.
 If you are interrupted as a developer (e.g., phone calls,
 urgent mail, someone asks a question) I don't think the
 first thing you do is to stop a timer on your computer.
 Moreover, developers often work on several defects in
 parallel, and might interweave bug fixing with new
 development. I don't think the "interruption time" captures
 all such multi-tasking, and it should at least be properly
 discussed in Section 7 "Threats to validity".

\begin{quote}{\em 
Yes, there is undoubtedly some accuracy in the time-tracking. We provide some evidence on the rigor of the TSP data collection process in \tion{defin}, \tion{data_integrity}, and \tion{data_character}, and a more extensive discussion on the issue of time-tracking and defect logging accuracy in \tion{conclusion_validity} on Conclusion Validity. We note that the data being collected here was not collected
for the purposes of this paper. Rather, it was being collected and used by the
programmers to assist them in their day-to-day management of the project.
Given this, we have confidence that the logging information is not widely inaccurate.

Please also consider the sanity checks reported in a companion paper (referenced here \tion{data_integrity}). In summary, this data has been heavily vetted.
}
\end{quote}
